\bf Title here

### 2020-08-19 paper/paper.Rmd
changes in introduction as per meeting 21

In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints. The problem consists in optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced space given by the problem constraints to either minimize or maximize a function.... Gradient based optimization has been typically used to solve such problems. However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem.

Derivative free methods provide another option to optimise the objective function without evaluating any gradient or Hessian information and a particular class of methods, direct search, has gained popularity through its conceptual simplicity.  However, the whole searching process in the algorithm remains a black-box. Plots are usually used to evaluate and compare the performance of different algorithms but it can easily become tedious because the code will have to be modified significantly when comparing different parameters in the algorithms. For example, a categorical variable with 5 levels can be easily mapped onto colour while mapping another categorical variable with 30 levels will not make the plot informative. Thus the plot needs to be re-designed to better suit the characteristics of the parameter (whether the parameter is a scalar or a matrix? whether the parameter is quantitative or categorical? If categorical, how many levels does the parameter have?). This motivates the  design of a visual diagnostic framework for optimisation algorithms based on the idea of a *global object*.

The remainder of the paper is organised as follows. Section \ref{DFO} gives a general overview of optimisation methods, specifically derivative free optimisation. 
Section \ref{tour} introduces the optimisation problem in a specific context: projection pursuit guided tour and existing derivative free algorithms for solving it. 
Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveils the black-box of optimisation. Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} gives definitions of different diagnostic plots using grammar of graphic and examples. 
Finally, Section \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.

# Derivative free optimisation {#DFO}

@@

## Motivation

We can see from Figure \ref{tour-path} that only the data projected using the accepted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation. Although these algorithms have been demonstrated to find the global maximum in various literature as we cited above, they focus solely on finding the global maximum and fewer tools are available to explore how each algorithm searches the parameter space and compare the results between different algorithms.

Visualisation has been widely used in exploratory data analysis. Presenting information in a graphical format often allows people to see information they would otherwise not see. This motivates us **to develop a visual framework to diagnose optimisation algorithms in projection pursuit with the aim to understand and compare different existing algorithms.** This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**.

@@

Hence the study **contributes to set up a tidy data structure for collecting observations from optimisation algorithms and prescribe a set of rules for generalising diagnostic plots for optimisation algorithms, specifically those used for projection pursuit.**

@@

# Visual diagnostic plots {#vis-diag-plots}

@@

# Implementation: Ferrn package {#implementation}

>>>>

<!-- We can see from Figure \ref{tour-path} that only the data projected using the accepted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation. Although these algorithms have been demonstrated to find the global maximum in various literature as we cited above, they focus solely on finding the global maximum and fewer tools are available to explore how each algorithm searches the parameter space and compare the results between different algorithms.  -->

<!-- This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**. -->

<!-- Hence the study **presents a gallery of diagnostic plots applicable to optimisation algorithms.** -->

Visualisation has been widely used in exploratory data analysis. Presenting information in a graphical format often allows people to see information they would otherwise not see. This motivates our work of creating plots to diagnose optimisation algorithms in the context of projection pursuit guided tour, with the aim to understand and compare features of different existing algorithms.

In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints. The problem consists in optimizing an objective function $f: S \rightarrow \mathbb{R}$ with $S \in \mathbb{R}^n$ in a reduced space given by the problem constraints to either minimize or maximize a function.

Projection pursuit and guided tour are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space. Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.

The remainder of the paper is organised as follows. 
Section \ref{optim} provides the literature review of optimisation methods, specifically the line search methods used in projection pursuit guided tour. 
Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem, and introduces three main existing algorithms.
Section \ref{vis-diag} presents the new visual diagnostics design, from forming the data object to the definition of different diagnostic plots with some small examples. 
Section \ref{application} shows the applicaiton of how the diagnostic plots designed in section \ref{vis-diag} can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.

# Derivative free optimisation {#optim}

@@

@@

@@

## Visual diagnostic plots {#vis-diag-plots}

@@

# Application {#application}
# Implementation: Ferrn package {#implementation}

### 2020-08-19 paper/paper.tex
changes in introduction as per meeting 21

In an optimization problem the goal is to find the best solution within
the space of all feasible solutions which typically is represented by a
set of constraints. The problem consists in optimizing an objective
function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
space given by the problem constraints to either minimize or maximize a
function\ldots{}. Gradient based optimization has been typically used to
solve such problems. However, there are many situations where
derivatives of an objective function are unavailable or unreliable and
therefore traditional methods based on derivatives are not the best
option to solve an optimization problem.

Derivative free methods provide another option to optimise the objective
function without evaluating any gradient or Hessian information and a
particular class of methods, direct search, has gained popularity
through its conceptual simplicity. However, the whole searching process
in the algorithm remains a black-box. Plots are usually used to evaluate
and compare the performance of different algorithms but it can easily
become tedious because the code will have to be modified significantly
when comparing different parameters in the algorithms. For example, a
categorical variable with 5 levels can be easily mapped onto colour
while mapping another categorical variable with 30 levels will not make
the plot informative. Thus the plot needs to be re-designed to better
suit the characteristics of the parameter (whether the parameter is a
scalar or a matrix? whether the parameter is quantitative or
categorical? If categorical, how many levels does the parameter have?).
This motivates the design of a visual diagnostic framework for
optimisation algorithms based on the idea of a \emph{global object}.

The remainder of the paper is organised as follows. Section \ref{DFO}
gives a general overview of optimisation methods, specifically
derivative free optimisation. Section \ref{tour} introduces the
optimisation problem in a specific context: projection pursuit guided
tour and existing derivative free algorithms for solving it. Section
\ref{vis-diag} and \ref{vis-diag-plots} together present the new
construction of a visual diagnostic framework that unveils the black-box
of optimisation. Section \ref{vis-diag} focuses on motivational ideas
and data structure and Section \ref{vis-diag-plots} gives definitions of
different diagnostic plots using grammar of graphic and examples.
Finally, Section \ref{implementation} mentions the R package: ferrn that
implements the visual diagnostics described above.

\hypertarget{DFO}{%
\section{Derivative free optimisation}\label{DFO}}

@@

\hypertarget{motivation}{%
\subsection{Motivation}\label{motivation}}

We can see from Figure \ref{tour-path} that only the data projected
using the accepted and interpolated bases are animated by tour; none of
the searching bases sampled by the optimsiation algorithms are presented
in the tour animation. Although these algorithms have been demonstrated
to find the global maximum in various literature as we cited above, they
focus solely on finding the global maximum and fewer tools are available
to explore how each algorithm searches the parameter space and compare
the results between different algorithms.

Visualisation has been widely used in exploratory data analysis.
Presenting information in a graphical format often allows people to see
information they would otherwise not see. This motivates us \textbf{to
develop a visual framework to diagnose optimisation algorithms in
projection pursuit with the aim to understand and compare different
existing algorithms.} This study is meaningful because \textbf{as more
complex index functions are proposed based on statistical theory, only
computationally effective optimisation algorithms allow us to find the
projection basis that presents interesting structure in the data}.

@@

Hence the study \textbf{contributes to set up a tidy data structure for
collecting observations from optimisation algorithms and prescribe a set
of rules for generalising diagnostic plots for optimisation algorithms,
specifically those used for projection pursuit.}

@@

\hypertarget{vis-diag-plots}{%
\section{Visual diagnostic plots}\label{vis-diag-plots}}

@@

>>>>

Visualisation has been widely used in exploratory data analysis.
Presenting information in a graphical format often allows people to see
information they would otherwise not see. This motivates our work of
creating plots to diagnose optimisation algorithms in the context of
projection pursuit guided tour, with the aim to understand and compare
features of different existing algorithms.

In an optimization problem the goal is to find the best solution within
the space of all feasible solutions which typically is represented by a
set of constraints. The problem consists in optimizing an objective
function \(f: S \rightarrow \mathbb{R}\) with \(S \in \mathbb{R}^n\) in
a reduced space given by the problem constraints to either minimize or
maximize a function.

Projection pursuit and guided tour are exploratory data analysis tools
that detect interesting structure of high dimensional data through
projection on low dimensional space. Optimisation is applied here to
search for the low dimensional space that finds the most interesting
projection.

The remainder of the paper is organised as follows. Section \ref{optim}
provides the literature review of optimisation methods, specifically the
line search methods used in projection pursuit guided tour. Section
\ref{tour} reviews projection pursuit guided tour, form the introduces
the optimisation problem, and introduces three main existing algorithms.
Section \ref{vis-diag} presents the new visual diagnostics design, from
forming the data object to the definition of different diagnostic plots.
Section \ref{application} shows some examples of how the diagnostic
plots designed in section \ref{vis-diag} can be applied to understand
and compare different algorithms and how they contribute to
modifications that improve the algorithms. Finally, Section
\ref{implementation} describes the R package: ferrn, that implements all
the visual diagnostics above.

\hypertarget{optim}{%
\section{Derivative free optimisation}\label{optim}}

@@

@@

@@

\hypertarget{vis-diag-plots}{%
\subsection{Visual diagnostic plots}\label{vis-diag-plots}}

@@

\hypertarget{application}{%
\section{Application}\label{application}}
